---
title: "Apply Combindation Test and Closed Testing Procedure for A Hypothetical Oncology Trial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Appling Combindation Test and Losed Testing Procedure for A Hypothetical Oncology Trial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1.  Background
Assume a sponsor is evaluating whether a new experimental drug, with the selected optimal dose, is superior to the control drug using an adaptive clinical trial. The primary endpoint is overall survival (OS). 

* The interim analysis (IA1) is planned for dose selection among three candidate dose levels of the experimental drug, compared to the control drug. IA1 occurs when 30 patients per arm have been enrolled. e.g., a total of 120 patients enrolled across three dose arms of the experimental drug and the control drug (a randomization ratio of 1:1:1:1).
*  After IA1, one dose of the experimental drug will be selected based on the pre-specified criteria. Additional patients will then be randomized in a 1:1 ratio between  the selected dose arm and the control arm until a total of 700 patients have been enrolled across these two arms.
*  IA2 and IA3 are interim analyses for OS, with the intention to stop early for efficacy. They will be triggered at 50% and 80% information fraction (IF), respectively.
 * The final analysis (FA) for OS is planned at 520 OS events. 

The statistical testing will be based on the combination test along with the closed testing procedure (Zhang et al. 2025). Patients enrolled before and after dose selection (IA1) will be combined using the inverse normal combination test. In total, there are three null hypotheses: dose 1 vs control (H01), dose 2 vs control (H02) and dose 3 vs control (H03), respectively. The null hypothesis of the selected dose arm will be rejected only if all possible intersection hypotheses that include the selected dose arm are rejected at the alpha level.

## 2.  Calculate OS efficacy boundary up to FA at the design satge 

First, let's set up the inputs. We begin by specifying the weights for Dunnett's test, based on incremental OS data from cohort 1 across IA2 to FA. At IA2, all the weights allocated to the data up to IA2, since this is the first stage for testing.    
```{r,warning = FALSE, message = FALSE}
library(gsDose)
w.ia2 <- c(1,0,0)
```

At IA3, we need to combine the data at IA2 with the incremental data accrued between IA2 and IA3. We set the weights proportional to the expected number of events at IA2, and the newly accrued events from IA2 to IA3. Similarly, at FA, incremental data across IA2, IA3 and FA need to be combined. We use gsDesign2 to estimate the expected number of events.
```{r,warning = FALSE, message = FALSE}
library(gsDesign2)

# design parameters
n1 <- 60
n2 <- 700
gamma <-  c(c( 5, 8, 12,18)*2/3,0,c(18,25),23)
R <- c(2,2,2,4,4,8,18,1)
lambda <- 33
hr <- 0.74
planE <- c(260,416,520)

# total enrollment for 700 #
enroll <- define_enroll_rate(
  duration = R,
  rate = gamma
)

# enrollment for cohort 1 60 pts #  
enroll.n1 <- define_enroll_rate(
  duration = c(2,2,2,2),
  rate = gamma[1:4]
)

# failure rate model
fail.rate <- define_fail_rate(
  duration = 1000,
  fail_rate = log(2) / lambda,
  hr = hr,
  dropout_rate = 0
)

# calculate the expected number of events from n1 at IA2/IA3/FA
l <- length(planE)
d.n1 <- rep(NA,l)

for(j in 1:l){
  
  # expected time for each IA
  exp.ia.time <- expected_time(
    enroll_rate = enroll, fail_rate = fail.rate,
    ratio = 1, target_event = planE[j],interval=c(0.01,1000))
  
  # expected number of events from cohort 1 at the expected time 
  d.n1[j] <- expected_event(
    enroll_rate = enroll.n1,
    fail_rate = fail.rate,
    total_duration = exp.ia.time$time, simple = TRUE)  
}

d21 <- d.n1[1]
d31 <- d.n1[2]
d41 <- d.n1[3]

# weights at IA3
w.ia3 <- c(sqrt(d21/d31),sqrt(1-d21/d31),0)
# weights at FA
w.fa <- c(sqrt(d21/d41),sqrt((d31-d21)/d41), sqrt((d41-d31)/d41) )

# set up the w matrix for cohort 1
w <- matrix(c(w.ia2,w.ia3,w.fa), nrow = 3, ncol = 3, byrow = TRUE)
colnames(w) <- c('IA2','IA3->IA4','IA3->FA')
row.names(w) <- c('IA2','IA3','FA')
w
```

We then create the h matrix of pre-specified weights for cohort 1 and cohort 2. We set them proportional to the number of events in each cohort. We also set the weights to be the same at each IA and FA.

```{r,warning = FALSE, message = FALSE}
h <- matrix(rep(c(sqrt(n1/n2),sqrt(1-n1/n2)),3),nrow = 3, ncol = 2, byrow = TRUE)
colnames(h) <- c('Cohort 1','Cohort2')
row.names(h) <- c('IA2','IA3','FA')
h
```

Next, we set the expected number of events cross IAs for cohort 1 and cohort 2, respectively. We also need to provide the planned number of events at FA
```{r,warning = FALSE, message = FALSE}
d <- planE 
d2 <- planE-ceiling(d.n1)
planD <- planE[3]
```
We then specify up to which stage the efficacy boundaries need to be calculated. Since FA is the third group sequential test in this trial, setting s = 3 indicates that the calculation should extend up to FA 
```{r,warning = FALSE, message = FALSE}
s <- 3
```
Lastly, we specify the total alpha. We use OBF alpha spending function to allocate alpha across IAs and FA. 
```{r,warning = FALSE, message = FALSE}
alpha <- 0.025
```
Okay, now let's run the function and generate the boutonnieres using gsBoundary. 
```{r,warning = FALSE, message = FALSE}
gsBoundary(w,h,d,d2,s,planD,alpha,sf='OF',bt='upper')

```
## 3.  Update efficacy boudnary and perform closed testing procedure based on the combination test

Let's say dose arm 2 is selected after IA1. IA2 occurs when 270 OS events have accrued, with 231 coming from cohort 2. We first update the boundary based on the actual number of events.
```{r,warning = FALSE, message = FALSE}
d.ia2 <- 270
d2.ia2 <- 231
s<-3
gsBoundary(w=1,h=h[1,],d.ia2,d2.ia2,s,planD,alpha,sf='OF',bt='upper')

```
Suppose we use the same weights w and h for all intersection hypotheses including individual hypothesis. The boundary at IA2 is 2.90 for all the hypotheses. Next, let's calculate the combination test statistics and perform the closed testing procedure. Assume the cohort 1 based standardized log-rank z statistics for each dose arm compared to the control arm at IA2 are 0.8, 1 and 0.7.  
```{r,warning = FALSE, message = FALSE}
z1.ia2 <- matrix(c(0.8,  1,   0.7),nrow = 1, ncol = 3, byrow = T)
colnames(z1.ia2) <- c('Dose 1 vs control','Dose 2 vs control','Dose 3 vs control')
row.names(z1.ia2) <- 'IA2'
z1.ia2
```
Suppose the number of events from cohort 1 at IA2 is 40 for Dose 1 vs. Control, 36 for Dose 2 vs. Control, and 38 for Dose 3 vs. Control, respectively.
```{r,warning = FALSE, message = FALSE}
d1.ia2 <- matrix(c(40, 36, 38),nrow = 1, ncol = 3, byrow = TRUE)
colnames(d1.ia2) <- c('Dose 1 vs control','Dose 2 vs control','Dose 3 vs control')
row.names(d1.ia2) <- 'IA2'
d1.ia2
```
Assume the standardized long-rank z statistic based on cohort 2 is 2.01 at IA2. Let's calculate the combination test and apply closed testing procedure. 
```{r,warning = FALSE, message = FALSE}
z2.ia2 <- 2.01
combTest(w=1,h=h[1,],z1.ia2,d1.ia2, z2.ia2, sid=2,zbound=2.9)
```
None of the intersection hypotheses that include H02 are rejected. As a result, the trial continues to IA3. Suppose IA3 is triggered by 430 events with 385 coming from cohort 2. The boundary at IA2 is 2.22.
```{r,warning = FALSE, message = FALSE}
d.ia3 <- c(270,430)
d2.ia3 <- c(231,385)
s<-3
gsBoundary(w=w[1:2,1:2],h=h[1:2,1:2],d.ia3,d2.ia3,s,planD,alpha,sf='OF',bt='upper')

```
Consider the cohort 1 based standardized log-rank z statistics for each dose arm compared to the control arm at IA3 are 0.9, 1.2 and 0.81; and the number of events from cohort 1 at IA3 is 51 for Dose 1 vs. Control, 42 for Dose 2 vs. Control, and 49 for Dose 3 vs. Control, respectively; and the standardized long-rank z statistic based on cohort 2 at IA3 is 2.35. 
```{r,warning = FALSE, message = FALSE}
z1.ia3 <- matrix(c(0.8,  1,   0.7,
                   0.9,  1.2, 0.81),nrow = 2, ncol = 3, byrow = TRUE)
colnames(z1.ia3) <- c('Dose 1 vs control','Dose 2 vs control','Dose 3 vs control')
row.names(z1.ia3) <- c('IA2','IA3')
z1.ia3

d1.ia3 <- matrix(c(40, 36, 38,
                   51, 42, 49),nrow = 2, ncol = 3, byrow = TRUE)
colnames(d1.ia3) <- c('Dose 1 vs control','Dose 2 vs control','Dose 3 vs control')
row.names(d1.ia3) <- c('IA2','IA3')
d1.ia3

z2.ia3 <- 2.35
combTest(w=w[2,1:2],h=h[2,],z1.ia3,d1.ia3, z2.ia3, sid=2,zbound=2.22)
```
All the intersection hypotheses that include H02 are rejected. Therefore, H02 is rejected and we claim dose arm 2 of the experimental drug is statistically better than the control drug in OS.  
